---
title: "Peer Assignment 1 - Reproducible Research"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Set the working directory where the CSV file is located.

```{r}
setwd("U:/R Directory/RepRes PA1")
```

Load the data

```{r}
library("foreign")
data<-read.csv("activity.csv", header=T)
attach(data)
summary(data)
```

## What is mean total number of steps taken per day?

Plot a histogram of the steps data, ignoring NAs

```{r}
DailySteps<-tapply(data$steps,data$date,FUN=sum,na.rm=T)
hist(DailySteps, breaks=20, col="red",xlab="Steps per Day", main="Histogram of Steps")
```

Calculate the mean and median number of steps per day

```{r echo=TRUE,results="hide"}
m1<-mean(DailySteps, na.rm=T)
med1<-median(DailySteps, na.rm=T)
```

The mean number of steps per day is `r m1`. 

The median number of steps per day is `r med1`.

## What is the average daily activity pattern?

Make a time series plot of the 5-minute intervals and the average number of steps taken, averaged across all days.

```{r echo=TRUE}
avg <- aggregate(x=data$steps,by=list(data$interval),FUN=mean, na.rm=TRUE)
attach(avg)
plot(Group.1,x,type='l',xlab='5-minute interval',ylab='Avg. steps taken')
maxint<-avg[which.max(avg$x),1]
```

The 5-minute interval, on average, with the most steps is `r maxint`.

## Imputing missing values

```{r}
NumberMissing<-sum(is.na(data$steps))
```

There are `r NumberMissing` missing values in the dataset.

Replace missing values with the mean for that 5-minute interval across days. The result is a new dataset, in this case called NewData

```{r}
ImpVals<-rbind(avg,avg,avg,avg,avg,avg,avg,avg,avg,avg,avg,avg)
ImpVals2<-rbind(ImpVals,ImpVals,ImpVals,ImpVals,ImpVals,avg)
NewData<-cbind(data,ImpVals2$x)
attach(NewData)
NewData$steps <- ifelse(is.na(steps),ImpVals$x,steps)
```

Make a histogram of the total number of steps taken each day. 

```{r}
DailySteps2<-tapply(NewData$steps,NewData$date,FUN=sum)
hist(DailySteps2, breaks=20, col="red",xlab="Steps per Day", main="Histogram of Steps")
```

Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE,results="hide"}
m2<-mean(DailySteps2)
med2<-median(DailySteps2)
```

The mean number of steps per day is `r m2`. 

The median number of steps per day is `r med2`.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor indicating if the given date is a weekday or weekend.

```{r}
NewData$date<-as.Date(NewData$date)
wkday<-weekdays(NewData$date)
NewData<-cbind(NewData,wkday)
NewData$daycat[wkday=="Monday"]<-"Weekday"
NewData$daycat[wkday=="Tuesday"]<-"Weekday"
NewData$daycat[wkday=="Wednesday"]<-"Weekday"
NewData$daycat[wkday=="Thursday"]<-"Weekday"
NewData$daycat[wkday=="Friday"]<-"Weekday"
NewData$daycat[wkday=="Saturday"]<-"Weekend"
NewData$daycat[wkday=="Sunday"]<-"Weekend"
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days.

Time series plot of the average steps taken on weekdays:

```{r}
steps.weekday <- aggregate(steps ~ interval, data = NewData, subset = NewData$daycat=="Weekday", FUN = mean)
plot(steps.weekday, type = "l")
```

Time series plot of the average steps taken on weekends:

```{r}
steps.weekend <- aggregate(steps ~ interval, data = NewData, subset = NewData$daycat=="Weekend", FUN = mean)
plot(steps.weekend, type = "l")
```
